\chapter*{Abstract}

Watching electronic sports (esports) has become a mainstream spectator activity. High-quality commentary can improve the viewing experience but is resource-intensive and unevenly available across competitive matches. This thesis investigates whether high-fidelity game telemetry can be transformed into engaging live text commentary for esports spectators. Using \textit{Age of Empires II: Definitive Edition} (AoE2) as a case study, and leveraging granular match data captured via \textit{CaptureAge}, we focus on generating incremental real-time match updates that convey progression in real time.

We first conduct a domain analysis that combines prior work on shoutcasting with a preliminary study, identifying viewer-relevant content and communication strategies for real-time reporting in this domain. Based on these findings, we present \textit{SAGE} (Semantic Abstraction \& Generation from Events), a modular data-to-text pipeline that converts high-fidelity match snapshots into reportable events and generates live text commentary using Large Language Models. Two engagement-oriented modules operationalize shoutcasting-inspired strategies: \textit{Expert Insights}, which adds strategic context and clarifies the significance of key moments, and \textit{Event Structuring}, which organizes events to improve narrative flow and readability.

We evaluate the prototype with an online user study of experienced AoE2 spectators (N=33) using a within-subjects ablation design across four high-level tournament matches. Participants rate linguistic quality and engageability using measures adapted from the Motivation Scale of Esports Spectatorship. Results show that \textit{Expert Insights} improves perceptions related to competitive nature, skill improvement, and vicarious sensation. \textit{Event Structuring} improves perceived logicality but does not, by itself, increase engageability for this audience, and we find no evidence of synergistic interaction effects between modules. Participants found the texts coherent and readable while reporting limited excitement, pointing to the challenges of engaging text-only reporting and the importance of macro-level context for interpreting complex real-time strategy play.

Together, these contributions provide an end-to-end, user-evaluated account of live data-to-text generation over a dense parallel event stream and inform the design of future automated commentary systems. We also release two new datasets of continuous game-state AoE2 snapshots, 173 tournament matches and 3,100 competitive ladder matches at 1-second and 60-second snapshot resolutions. This set covers entity positions, player economy, technology state, fog of war, map layout snapshots, and layout, event and command streams.

\vfill
\begin{keywords}
  esports; data-to-text generation; natural language generation; live commentary; Age of Empires II; large language models; user study; dataset; game telemetry
\end{keywords}
