\chapter*{Acknowledgements}\label{chap:personal}
This work represents the culmination of a six-year journey (enough time to be skeptical of (not so) Large Language Models and then embrace them) that would not have been possible without the support of many incredible people. While it is difficult to fully capture the depth of my gratitude in words, I want to take this opportunity to express my appreciation.

First and foremost, I would like to thank my supervisors. \textbf{Mariët}, thank you for your unwavering support and guidance throughout this process, and for the many times you went above and beyond to help me get back up and running after periods away from this research. Your encouragement, your belief in my work, and your endless ways of keeping me motivated have been a constant source of inspiration. \textbf{Lorenzo}, thank you for your insightful feedback and your invaluable technical and practical guidance, which truly shaped the direction of this project, and for finding many clever ways to motivate me and steer my creativity another way. \textbf{Vadim}, thank you for your unique perspective and for providing the critical feedback to push this work further. But perhaps the most valuable aspect of their guidance was sharing their experience as researchers; they explained to me how researchers work, what they value, and how that should be reflected in my work to ensure a meaningful and useful contribution to the field.

To my family and friends: \textbf{Mom}, \textbf{Dad}, \textbf{Sister}, and \textbf{Brother-in-law}, thank you for your unconditional love and support, especially for standing alongside me during the most challenging times. In particular, I am deeply grateful to my father, who spent many nights trying to understand the full breadth and depth of this work, and to my mother, who was never short of emotional support. Your encouragement has been a constant source of strength (and, from time to time, a healthy dose of frustration). A special thanks to my friends \textbf{Midas} (Thomas) and \textbf{Lieke} for their enthusiasm, their support, and for truly understanding my drive to see this through to the very end. I must also thank my dog, \textbf{Mara}, who was never short of enthusiasm herself, always offering the perfect support in the form of a timely walk or a cuddle.

The groundwork for this thesis was laid while I was working full-time at \textbf{CaptureAge}, which was a truly life-altering rollercoaster; an experience that led me on an unbelievable adventure alongside some of the most technically talented people I have ever had the pleasure of working with. Being able to leverage the technology we built at CaptureAge for this research has been an immense privilege. I am grateful to the company for allowing me to do so and for supporting the publication of the derived dataset. 

I also want to extend my gratitude to the competitive \textbf{Age of Empires II community}. Thank you for your enthusiastic participation in both studies, and for providing such honest and direct feedback; the only real type of feedback. May your Wololos be quick, and your villagers safe from boars.

Additionally, I would like to thank \textbf{World's Edge} for permitting the use of Age of Empires II and its recorded games, as well as the publication of the derived data. My thanks go to \textbf{MembTV} for allowing me to use his King of the Desert IV matches for the evaluation, and to \textbf{MembTV and Chris}, \textbf{T90Official}, \textbf{LidaKoR}, and \textbf{NiliAoE} for allowing me to use their shoutcasts as a data source. Finally, a special thank you to \textbf{DonDinardoni} for his invaluable help in obtaining and pre-processing the replays.