\chapter{Generated Outputs used for Evaluation}
\label{app:generated-outputs}
\section{Vinchester (IronBoar) vs MbL (TC\_Eater)}
\label{appendix:vinchester-mbl}
\subsection{Base}

\textbf{IronBoar (Britons, yellow) vs TC\_Eater (Huns, red)}

0:04 - Welcome to Arabia! In the yellow corner, playing as the Britons, we have IronBoar. Facing off against them, in the red corner, it's TC\_Eater, leading the Huns! \par \vspace{1em}

0:04 - IronBoar has a very strong start with main gold and stone near the TC, and excellent nearby woodlines. TC\_Eater also finds their main gold and stone quite close, with multiple accessible woodlines. Both players have solid home maps! \par \vspace{1em}

11:06 - We've just seen a raid near IronBoar's base! TC\_Eater's Scout and Militia pushed in, taking down IronBoar's Scout and two Militia. TC\_Eater only lost one Militia in return! TC\_Eater is now ahead in villager count with 25 to IronBoar's 24. \par \vspace{1em}

15:21 - A huge battle at TC\_Eater's main gold mine! IronBoar's Archers and Skirmishers clashed with TC\_Eater's units. IronBoar lost 1 Skirmisher, but took down 4 Skirmishers and a Scout Cavalry from TC\_Eater! What a win for IronBoar! \par \vspace{1em}

16:19 - A brutal skirmish near TC\_Eater's base! IronBoar lost 3 Archers for just 1 Lumberjack from TC\_Eater. IronBoar is now pushing for Castle Age, but TC\_Eater continues to mass Skirmishers, now with 10 total! \par \vspace{1em}

16:52 - Another skirmish on the left hill! IronBoar lost an Archer, TC\_Eater a Skirmisher. IronBoar is now heading to Castle Age! Meanwhile, TC\_Eater is still Feudal, researching Padded Archer Armor and has 10 Skirmishers. \par \vspace{1em}

18:16 - TC\_Eater has completed a Watch Tower! It's up near their main Gold Mine, looking to secure that vital resource as they continue to mass units. \par \vspace{1em}

18:41 - A standoff at IronBoar's forward gold! TC\_Eater's Skirmishers and Archers confronted IronBoar's Builders and Scout Cavalry. No units were lost by either side! IronBoar is still researching Castle Age, while TC\_Eater is researching Wheelbarrow with a strong military force. \par \vspace{1em}

18:45 - Big battle at IronBoar's stone mine! IronBoar lost 2 units, but took down 3 Skirmishers from TC\_Eater. IronBoar is still pushing Castle Age. TC\_Eater is now at 40 villagers and continues military production! \par \vspace{1em}

21:09 - Huge battle near IronBoar's base! IronBoar lost 2 Skirmishers but wiped out ALL 11 of TC\_Eater's units! IronBoar is now in Castle Age and making Crossbowmen. TC\_Eater is still researching Castle Age. \par \vspace{1em}

21:16 - Another battle near TC\_Eater's base! IronBoar's Crossbowmen destroyed three of TC\_Eater's Crossbowmen, losing nothing! IronBoar is firmly in Castle Age, now with a second TC. TC\_Eater is still researching Castle Age but holds the villager lead! \par \vspace{1em}

22:26 - Heavy losses for IronBoar near TC\_Eater's base! IronBoar's Crossbowmen lost 2 units, while TC\_Eater took no casualties. TC\_Eater still leads in villagers, but IronBoar has its second Town Center up! \par \vspace{1em}

26:28 - A massive raid at TC\_Eater's base! IronBoar lost 10 units, but TC\_Eater lost 15, including 7 villagers! IronBoar is now at 50 villagers, TC\_Eater at 41. Both are in Castle Age, but TC\_Eater still leads in army size! \par \vspace{1em}

27:02 - A brutal engagement at TC\_Eater's stone mine! IronBoar lost all 3 of its units, including a Knight, while TC\_Eater took no casualties! TC\_Eater reinforces its army, now with 32 military units! \par \vspace{1em}

28:36 - A tense standoff in the south as IronBoar's Crossbowmen faced TC\_Eater's archers! No losses for either side. IronBoar booms with a third Town Center and adds a Mangonel. TC\_Eater researches archer armor, still leading in military size. \par \vspace{1em}

31:33 - IronBoar's Knights were crushed on TC\_Eater's hill, losing all three! But they took 3 Lumberjacks with them. TC\_Eater still leads in military units, IronBoar in villagers! \par \vspace{1em}

31:45 - A massive battle on IronBoar's forward hill! IronBoar lost 12 units, but absolutely crushed TC\_Eater, who lost 33 units, including most of their Elite Skirmishers and Crossbowmen! What a trade for IronBoar! \par \vspace{1em}

34:18 - A skirmish in the west! IronBoar lost a Knight, while TC\_Eater lost a Crossbowman. IronBoar is now researching Imperial Age and built a fifth Town Center! TC\_Eater continues to mass military, now with 43 units. \par \vspace{1em}

37:41 - Massive battle near IronBoar's base! IronBoar lost 13 units, but TC\_Eater was decimated, losing ALL 32 units including their Elite Skirmishers and Crossbowmen! IronBoar is now Imperial Age! \par \vspace{1em}

39:37 - A massive engagement in the west! IronBoar's Arbalesters and Mangonel took on TC\_Eater's Elite Skirmishers and Crossbowmen. IronBoar lost 7 units but wiped out 20 of TC\_Eater's! IronBoar's Imperial Age push looks strong! \par \vspace{1em}

40:43 - And TC\_Eater finishes a new Town Center in the south-west! A crucial economic boost as they try to recover from those recent devastating battles. They need this as IronBoar is already Imperial! \par \vspace{1em}

41:46 - IronBoar has completed a Castle! This major fortification is up in the west, adding incredible power to their Imperial Age forces! What a statement! \par \vspace{1em}

41:58 - TC\_Eater has completed a Watch Tower! It's up near their forward gold mine, crucial for securing those valuable resources and fending off pressure. \par \vspace{1em}

42:02 - TC\_Eater has completed another Watch Tower! This one is up near their base, providing crucial defensive vision as IronBoar continues to push with Imperial Age power! \par \vspace{1em}

42:07 - Two massive battles unfold! Near IronBoar's base, they lost 4 units but took down 10 of TC\_Eater's, including Rams! Simultaneously, TC\_Eater raided IronBoar's west, killing 9 villagers but losing 15 military units! \par \vspace{1em}

42:08 - And that's it! The game has ended! IronBoar is the winner! Congratulations to IronBoar for taking the victory! \par \vspace{1em}

\subsection{Expert Insights}
\textbf{IronBoar (Britons, yellow) vs TC\_Eater (Huns, red)}

0:04 - Welcome to Arabia! In yellow, we have IronBoar, known for aggressive plays, commanding the Britons, masters of ranged units with their fast shepherds. Facing him in red is TC\_Eater, a defensive player, leading the Huns, a powerful cavalry civ strong at rushing, despite their initial wood deficit. Let's go! \par \vspace{1em}

0:04 - IronBoar's Britons have a very forward main gold and stone, which will demand early walling! TC\_Eater's Huns also face forward resources. This map setup could force both players into early aggression.IronBoar's Britons have a very forward main gold and stone, which will demand early walling! TC\_Eater's Huns also face forward resources. This map setup could force both players into early aggression. \par \vspace{1em}

7:04 - It's confirmed! TC\_Eater is going for a Drush! He's rushing his opponent with early militia in the Dark Age. This could put massive pressure on IronBoar and disrupt his economy right away! \par \vspace{1em}

9:04 - Unbelievable! IronBoar is also going for a Drush! We have a rare double Drush on our hands! Both players sending early militia. This will be an incredibly aggressive Dark Age! \par \vspace{1em}

11:04 - Confirmed! IronBoar is transitioning into archers for his Britons! This is a powerful play, leveraging his civilization's strengths. Expect those longbows to hit hard! \par \vspace{1em}

11:06 - A raid near IronBoar's base has concluded! IronBoar lost 3 units, while TC\_Eater lost just 1 in that engagement. IronBoar is now fielding archers and TC\_Eater continues with militia. Both are in Feudal Age! \par \vspace{1em}

15:21 - IronBoar's archers launched a devastating attack on TC\_Eater's main gold mine! TC\_Eater lost 4 Skirmishers and a Scout Cavalry, while IronBoar only lost 1 Skirmisher. Those Fletching and Padded Archer Armor upgrades are really paying off for IronBoar! \par \vspace{1em}

16:19 - IronBoar's archers just took heavy losses in a new skirmish near TC\_Eater's base, losing 3 archers to just a lumberjack! But wait, IronBoar is clicking up to Castle Age! TC\_Eater is also upgrading his Skirmishers with Padded Archer Armor. \par \vspace{1em}

16:52 - A skirmish just ended! IronBoar lost an Archer, while TC\_Eater lost a Skirmisher. IronBoar is now researching Castle Age, but TC\_Eater has ramped up his military, fielding more skirmishers in Feudal! \par \vspace{1em}

18:16 - TC\_Eater has completed a Watch Tower at his main gold! A smart defensive move, this will protect his vital resource from IronBoar's archer pressure! \par \vspace{1em}

18:41 - A battle at IronBoar's main gold mine ended with no losses! IronBoar is still researching Castle Age. Meanwhile, TC\_Eater, with more military in Feudal, is now researching Wheelbarrow, boosting his economy. \par \vspace{1em}

18:45 - A major battle just concluded at IronBoar's stone mine! IronBoar's archers defeated TC\_Eater's skirmishers, losing 2 units to TC\_Eater's 3. IronBoar is still clicking to Castle Age as TC\_Eater continues to build his economy. \par \vspace{1em}

21:09 - Absolute dominance! IronBoar's Crossbowmen wiped out TC\_Eater's forces near IronBoar's base, costing TC\_Eater 11 units to IronBoar's 2! IronBoar is now fully established in Castle Age with Bodkin Arrow! \par \vspace{1em}

21:16 - IronBoar's Crossbowmen are relentless! They just won a huge fight near TC\_Eater's base, taking zero losses while TC\_Eater lost 3 Crossbowmen. IronBoar's military lead and score are soaring! TC\_Eater is now researching Castle Age. \par \vspace{1em}

22:26 - TC\_Eater's defense pays off! In a battle near his base, IronBoar lost 2 Crossbowmen, but TC\_Eater took no casualties. TC\_Eater still pushing for Castle Age! \par \vspace{1em}

26:28 - IronBoar's aggressive raid just rocked TC\_Eater's base! IronBoar lost 10 military units, but TC\_Eater suffered even more, losing 8 military units and 7 vital villagers. Massive economic damage! \par \vspace{1em}

27:02 - Huge win for TC\_Eater! He decisively defended his stone mine, losing nothing while IronBoar lost 3 units. TC\_Eater's military strength is rising! \par \vspace{1em}

28:36 - A southern skirmish concluded with no losses for either side! However, IronBoar has lost a Town Center! Meanwhile, TC\_Eater's military is growing, and he now has Leather Archer Armor for his Elite Skirmishers. \par \vspace{1em}

31:33 - A massive clash! IronBoar's 3 Knights were annihilated at TC\_Eater's forward position, though TC\_Eater lost 3 crucial Lumberjacks. IronBoar's Knight aggression was repelled, but TC\_Eater took economic damage! \par \vspace{1em}

31:45 - What a colossal clash at IronBoar's forward! IronBoar's Knights, Crossbowmen, and Mangonels just annihilated TC\_Eater's army. TC\_Eater lost a staggering 33 units to IronBoar's 12! A decisive victory for IronBoar! \par \vspace{1em}

34:18 - A western skirmish concludes! IronBoar loses a Knight, but TC\_Eater also drops a Crossbowman in the exchange. IronBoar is now pushing for Imperial Age, holding a villager lead over TC\_Eater's larger Castle Age army! \par \vspace{1em}

37:41 - IronBoar crushed TC\_Eater near his base! TC\_Eater lost 32 units to IronBoar's 13 in a decisive victory. IronBoar is now in Imperial Age, leaving TC\_Eater behind in Castle! \par \vspace{1em}

39:37 - IronBoar's Imperial Arbalesters and Mangonel decimated TC\_Eater's Castle Age army in the west! TC\_Eater lost 20 units, while IronBoar only lost 7. IronBoar is clearly in control! \par \vspace{1em}

40:43 - TC\_Eater has finished a new Town Center in the south-west! A crucial economic boost for TC\_Eater, who is still in Castle Age against IronBoar's Imperial push. He's trying to boom back into this game! \par \vspace{1em}

41:46 - IronBoar drops a new Castle in the west! This powerful structure gives his Britons crucial map control and access to their unique units, pushing his advantage even further! \par \vspace{1em}

41:58 - TC\_Eater has completed a Watch Tower at his forward, tertiary gold mine! A defensive play to protect a vulnerable resource from IronBoar's pressure! \par \vspace{1em}

42:02 - TC\_Eater adds another defensive Watch Tower near his base! He's trying to fortify his economy against IronBoar's relentless Imperial Age push. Every defense counts now! \par \vspace{1em}

42:07 - TC\_Eater's siege push failed at IronBoar's base, losing 10 units to IronBoar's 4. Another western raid saw TC\_Eater lose 15 units to IronBoar's 13. IronBoar retains Imperial Age control! \par \vspace{1em}

42:08 - That's it! The game has ended! IronBoar has won! What an incredible performance, dominating with those Imperial Age Arbalesters and securing a decisive victory! \par \vspace{1em}

\subsection{Event Structuring}
\textbf{IronBoar (Britons, yellow) vs TC\_Eater (Huns, red)}

0:04 - Welcome to Arabia! Yellow IronBoar, playing as the Britons, faces off against Red TC\_Eater, commanding the Huns. Both players begin with their standard starting villagers and scouts, ready to explore this open map. Who will seize the early advantage? \par \vspace{1em}

0:04 - Yellow IronBoar has a forward main gold and main stone, both on the left. Multiple forward woodlines are also on the left. For Red TC\_Eater, his main gold is forward at the bottom of his base, with his main stone forward to his top-left. His woodlines are forward, scattered left, center, and bottom-right. \par \vspace{1em}

11:06 - A fierce raid just concluded near IronBoar's base! TC\_Eater's Scout Cavalry and three Militia struck, clashing with IronBoar's defending villagers, lumberjacks, a Scout, and three Militia. IronBoar takes significant casualties, losing one Scout Cavalry and two Militia, while TC\_Eater only loses a single Militia. TC\_Eater has now advanced to Feudal Age, and IronBoar has pushed two Archers forward! \par \vspace{1em}

15:21 - Big fights on the map! A major battle just finished at TC\_Eater's forward gold mine, where IronBoar's Archers and Skirmishers faced off against TC\_Eater's Skirmishers and Scout Cavalry. IronBoar emerged victorious, losing only one Skirmisher, while TC\_Eater lost four Skirmishers and a Scout Cavalry! Now, IronBoar's Archers are actively raiding near TC\_Eater's base, targeting villagers. IronBoar maintains a slightly larger military presence. \par \vspace{1em}

18:45 - IronBoar's Archers are relentless! They've just wiped out TC\_Eater's Skirmishers near his base and on a neutral hill, though IronBoar lost some Archers in the process. TC\_Eater is now reacting, completing a Watch Tower at his forward gold mine to bolster defenses. Meanwhile, IronBoar has started researching Castle Age! TC\_Eater holds a villager and military lead as the game progresses. \par \vspace{1em}

22:26 - What a turn of events! IronBoar's defense near his base was brutal, his Crossbowmen, Knights, and Scouts devastating TC\_Eater's Skirmishers and Crossbowmen, who lost 11 units. Meanwhile, IronBoar's Crossbowmen continue to raid TC\_Eater's base. IronBoar has now hit Castle Age and built a second Town Center, pulling ahead in score and military. TC\_Eater is still in Feudal, though researching Castle Age. \par \vspace{1em}

28:36 - IronBoar's aggressive raid at TC\_Eater's base traded military for several crucial villagers! But then, TC\_Eater's Elite Skirmishers wiped out IronBoar's attack at his forward stone. Now, a massive battle rages at IronBoar's forward hill, with his Knights and Crossbowmen defending a Town Center against TC\_Eater's strong skirmisher and crossbow force! TC\_Eater has the military advantage, while IronBoar has significantly more villagers. \par \vspace{1em}

31:45 - The massive battle at IronBoar's forward hill has just concluded! IronBoar's Town Center, Knights, and Mangonels successfully defended against a huge push from TC\_Eater's Skirmishers and Crossbowmen. TC\_Eater lost a staggering 33 units in that exchange, while IronBoar lost 12! A hard-fought defensive victory for IronBoar, but both sides took heavy military losses. \par \vspace{1em}

34:18 - A skirmish just wrapped up in the west, where TC\_Eater's Elite Skirmishers and Crossbowmen took down IronBoar's Knight, losing one Crossbowman in return. Meanwhile, the massive battle near IronBoar's base continues to rage! IronBoar's Crossbowmen, Mangonel, and Knight are holding off TC\_Eater's large force of Elite Skirmishers and Crossbowmen. IronBoar is researching Imperial Age, while TC\_Eater is still in Castle Age. IronBoar maintains a strong villager lead, but TC\_Eater has the larger military. \par \vspace{1em}

42:08 - IronBoar's defenses were unyielding! He inflicted massive losses on TC\_Eater's pushes, wiping out dozens of Skirmishers and Crossbowmen in major battles near his base and in the west, while losing far fewer units. IronBoar even successfully defended against Battering Rams! With IronBoar now in Imperial Age and building a Castle, TC\_Eater's attempts to push ran out of steam. IronBoar secures the victory! \par \vspace{1em}

\subsection{Expert Insights \& Event Structuring}
\textbf{IronBoar (Britons, yellow) vs TC\_Eater (Huns, red)}

0:04 - Welcome to Arabia! In yellow, it's IronBoar, known for his aggressive play, leading the Britons with their powerful archers and shepherd bonus. Across the map, we have TC\_Eater, a master of laming strategies, in red, commanding the Huns, a strong cavalry civ that starts with max pop cap. Let the battle begin! \par \vspace{1em}

0:04 - Both players have very forward main gold mines on this Arabia map, setting up for early contention! IronBoar, as Britons, will need to quickly secure those resources. TC\_Eater's Huns also face exposed gold, which could force an early aggressive push. \par \vspace{1em}

11:06 - TC\_Eater launched an early Drush, pushing aggressively with Militia! We just saw a skirmish near IronBoar's base, where TC\_Eater’s Militia caused chaos, taking down a Scout Cavalry and two Militia from IronBoar, only losing one of his own. IronBoar, initially also attempting a Drush, is now pivoting to Archers in the Feudal Age, while TC\_Eater leads in score and villager count. \par \vspace{1em}

15:21 - IronBoar strikes hard! His Archers and Skirmishers just routed TC\_Eater's defense at his forward gold, taking down 5 units including Skirmishers and a Scout for only one Skirmisher lost. IronBoar's Archer pressure is relentless, now with Fletching and Padded Archer Armor! He's currently engaging a lone Villager near TC\_Eater's base. TC\_Eater has more villagers but fewer military units now! \par \vspace{1em}

18:45 - IronBoar is making a crucial move, initiating his Castle Age research! He also successfully defended his forward secondary Stone Mine, losing 2 units but taking down 3 of TC\_Eater's Skirmishers. Meanwhile, TC\_Eater solidified his forward gold with a new Watch Tower. TC\_Eater still holds a strong villager lead with 40 to IronBoar's 33, but IronBoar's age-up could be a game changer! \par \vspace{1em}

22:26 - IronBoar is now firmly in the Castle Age, with Bodkin Arrow and Crossbowmen online! He completely wiped out TC\_Eater's Feudal army at his base and pushed forward aggressively. However, TC\_Eater is now researching Castle Age and defending well. His Skirmishers, still Feudal, are managing to hold off IronBoar's Crossbowmen near TC\_Eater's base, costing IronBoar two Crossbowmen! TC\_Eater maintains his villager lead. \par \vspace{1em}

28:36 - IronBoar's Castle Age aggression is relentless! He raided TC\_Eater's base, taking down 15 units from TC\_Eater, but lost 10 of his own in the process. TC\_Eater quickly responded with Elite Skirmishers, completely routing IronBoar's attack at his forward stone mine, costing IronBoar 3 units for free! Now, a major battle is brewing at IronBoar's forward hill! IronBoar leads in villager count, but TC\_Eater commands a significantly larger army. \par \vspace{1em}

31:45 - The major battle at IronBoar's forward hill has concluded with a decisive victory for IronBoar! His Crossbowmen, Knights, and Mangonels absolutely decimated TC\_Eater's Elite Skirmishers and Crossbowmen. TC\_Eater lost 33 units, while IronBoar lost 12. Earlier, TC\_Eater's Knight push at his own forward hill was also repelled. A small raid continues near IronBoar's base, but IronBoar now holds a commanding villager lead and superior army value. \par \vspace{1em}

34:18 - IronBoar is charging towards Imperial Age, a massive power spike! However, TC\_Eater is not letting up, still in Castle Age but with a huge army of Elite Skirmishers and Crossbowmen! A small skirmish saw IronBoar lose a Knight. Crucially, a full-blown battle is now raging right at IronBoar's base, with TC\_Eater’s military significantly outnumbering IronBoar’s! This Imperial Age push is in serious danger! \par \vspace{1em}

42:08 - What a defense from IronBoar! The huge battle at his base concluded with IronBoar's Imperial Age Arbalesters and Mangonels routing TC\_Eater, destroying 32 of his units for only 13 losses! IronBoar even completed a Castle and a new Town Center, while TC\_Eater added defensive towers. Despite TC\_Eater's continued pressure and persistent raids, IronBoar has held on and emerged victorious! IronBoar wins! \par \vspace{1em}

\section{Yo (FeudalLord) vs DauT (LoomRush)}
\label{appendix:yo-daut}
\subsection{Base}
\textbf{FeudalLord (Aztecs, blue) vs LoomRush (Franks, red)}

0:05 - Welcome to Arabia! We've got a fantastic matchup underway as Blue FeudalLord, leading the Aztecs, faces off against Red LoomRush, commanding the Franks. Both players are just beginning to explore their starting positions and gather initial resources. It's set to be a classic clash! \par \vspace{1em}

0:05 - Both players seem to have good starting maps! Blue FeudalLord has main gold and stone quite close, with accessible woodlines. Red LoomRush also has their main gold and stone nearby, with solid woodlines available. \par \vspace{1em}

10:29 - And just like that, the action begins! Blue FeudalLord's Man-at-Arms attack Red LoomRush's base, near the woodline. They take down a villager! A successful early raid for FeudalLord, drawing first blood! \par \vspace{1em}

11:59 - LoomRush's Scout Cavalry has completely wiped out FeudalLord's Man-at-Arms near his forward position! All three of Blue's raiders are gone, with Red losing no units. A big swing in military momentum! \par \vspace{1em}

15:06 - Blue FeudalLord's offensive push near Red LoomRush's base ends in disaster! All 5 of FeudalLord's Archers and Spearmen are wiped out. LoomRush's Skirmishers and Scout Cavalry dominate, losing only one Scout. A massive military victory for Red! \par \vspace{1em}

20:46 - Brutal battle near FeudalLord's base! Both players lost 7 units. Crucially, LoomRush is already in Castle Age, producing Knights! FeudalLord is still researching. A big tech lead for Red! \par \vspace{1em}

25:59 - A huge battle near FeudalLord's base! Red LoomRush's Knights and Cavalry Archers dominate, taking down Blue's Eagle Scouts and villagers, losing only two Skirmishers. Blue FeudalLord is now Castle Age, expanding to three Town Centers and ramping up Pikemen! \par \vspace{1em}

28:55 - Another brutal battle near FeudalLord's base! Red LoomRush's Knights and Cavalry Archers prove dominant, losing only two units while FeudalLord loses seven, including Monks. LoomRush maintains military pressure! \par \vspace{1em}

32:00 - FeudalLord is booming, adding a FOURTH Town Center and rapidly expanding their economy! Both players continue to mass military units. LoomRush maintains a strong army, while FeudalLord's forces are also growing. \par \vspace{1em}

32:27 - A massive battle erupts at FeudalLord's forward stone mine! LoomRush's Knights and Cavalry Archers crush FeudalLord's forces, inflicting heavy losses of 27 units while losing only 9. Despite the defeat, FeudalLord continues booming with more villagers! \par \vspace{1em}

32:31 - A skirmish erupts near LoomRush's base! FeudalLord loses a Knight, but takes down a LoomRush villager. Meanwhile, FeudalLord continues booming with four Town Centers, while LoomRush only has one! \par \vspace{1em}

35:16 - A battle at FeudalLord's base sees Blue lose an Eagle Warrior! But LoomRush loses three Cavalry Archers. FeudalLord continues to expand his economy, now at 73 villagers! \par \vspace{1em}

37:59 - Blue FeudalLord completes a Castle! It's a key forward position for Blue, built near a secondary stone mine. \par \vspace{1em}

36:48 - Massive clash near LoomRush's base! FeudalLord's Eagle Warriors and Pikemen are wiped out, losing 30 units. LoomRush loses only 10 Knights and Man-at-Arms. A decisive victory for Red LoomRush! \par \vspace{1em}

38:28 - Red LoomRush completes a new Town Center! It's built at a tertiary gold mine, a key economic expansion for LoomRush, boosting their villager production! \par \vspace{1em}

37:24 - A brutal major battle near LoomRush's Town Center! FeudalLord's army is decimated, losing 27 units, while LoomRush loses only 14, mostly villagers. LoomRush still holds a strong military advantage! \par \vspace{1em}

39:26 - Blue FeudalLord has completed another Castle! This one is defensively placed right near FeudalLord's starting Town Center. A secure position for Blue! \par \vspace{1em}

38:45 - Red LoomRush's Knights and Cavalry Archers strike at FeudalLord's forward stone mine! Blue loses a builder, but LoomRush takes no losses. Red maintains their military advantage! \par \vspace{1em}

40:13 - A brutal major battle erupts at LoomRush's Town Center! FeudalLord loses all 13 Eagle Warriors. LoomRush loses 14 units, including 8 villagers and 5 Long Swordsmen. A costly engagement for both! \par \vspace{1em}

41:26 - A massive battle near FeudalLord's base! Blue loses 30 units, including Monks and Eagle Warriors, while Red loses 39 units, mostly Long Swordsmen and Knights. Blue is now researching Imperial Age! \par \vspace{1em}

41:55 - Another brutal major battle at LoomRush's forward gold mine! FeudalLord's 8 Eagle Warriors are wiped out. LoomRush loses 4 Long Swordsmen and 3 villagers but holds. FeudalLord is now researching Imperial Age! \par \vspace{1em}

43:12 - A skirmish in the west! Blue FeudalLord loses a Pikeman, but Red LoomRush's Long Swordsmen completely dominate, taking no losses. FeudalLord is still researching Imperial Age! \par \vspace{1em}

44:16 - A major battle in the west! FeudalLord's Jaguar Warriors and Eagles crush LoomRush's Cavalry Archers and Long Swordsmen. LoomRush lost 16 units to FeudalLord's 5. Blue is still researching Imperial! \par \vspace{1em}

44:16 - AND THE GAME IS OVER! Blue FeudalLord has defeated Red LoomRush! What a comeback by FeudalLord, securing the victory on Arabia! Congratulations to FeudalLord! \par \vspace{1em}

\subsection{Expert Insights}
\textbf{FeudalLord (Aztecs, blue) vs LoomRush (Franks, red)}

0:05 - Welcome to Arabia! In blue, FeudalLord, known for his clever strategies, commands the Aztecs, a civilization focused on fast military production. In the red corner, LoomRush, renowned for his strategic depth, leads the Franks, a cavalry powerhouse with strong Knights and economic bonuses. \par \vspace{1em}

0:05 - FeudalLord has a very forward main gold, close to his TC, favoring aggressive Aztec play. LoomRush also sees a forward main gold and stone, setting up for early Frankish Knights. Both players will need early defense! \par \vspace{1em}

8:05 - FeudalLord is going for a Drush! This early aggression from the Aztec player aims to disrupt LoomRush's economy. A bold move from FeudalLord; let's see how LoomRush reacts to this pressure! \par \vspace{1em}

10:05 - Confirmation! FeudalLord is indeed going for a Man-at-Arms Rush. This early infantry aggression from the Aztecs will surely test LoomRush's defenses right out of the gate! \par \vspace{1em}

10:05 - LoomRush, the Frankish player, is now pivoting to Scouts! This could be a counter-move against the Man-at-Arms or an early raid attempt by LoomRush. What a response! \par \vspace{1em}

10:29 - FeudalLord's Man-at-Arms raid paid off! They've secured the first villager kill against LoomRush, a crucial hit to his economy. Meanwhile, LoomRush's Scouts are out, but FeudalLord's aggressive infantry is already making an impact! \par \vspace{1em}

13:05 - FeudalLord is now transitioning into ARCHERS! This is a smart pivot for the Aztec player, potentially countering LoomRush's earlier Scouts. What a strategic shift! \par \vspace{1em}

11:59 - LoomRush's Scouts have decisively crushed FeudalLord's Man-at-Arms! All three M@A are lost for FeudalLord, completely shutting down that early aggression. LoomRush's defense holds strong, leaving FeudalLord to rely on those incoming Archers! \par \vspace{1em}

15:06 - What a slaughter! LoomRush's Scouts and Skirmishers just crushed FeudalLord's Archers and Spearmen, losing only one Scout! FeudalLord lost his entire attacking force. LoomRush is now building a Market, showing his economic strength! \par \vspace{1em}

20:46 - LoomRush's Castle Age Knights just decimated FeudalLord's forces near his base! In a brutal 7-7 unit trade, FeudalLord lost his Archers and Eagle Scouts. LoomRush now holds the clear age advantage! \par \vspace{1em}

25:59 - LoomRush’s Knights and Archers just decimated FeudalLord’s Eagle Scouts and Pikemen on a forward hill! LoomRush took minimal losses, while FeudalLord lost key military and builders. FeudalLord is booming TCs, while LoomRush adds more powerful Knights! \par \vspace{1em}

28:55 - Another devastating blow for FeudalLord! LoomRush's Knights and Cavalry Archers just crushed FeudalLord's defending Pikemen, Monks, and Cavalry Archers near his base! FeudalLord lost seven units, while LoomRush only lost two. This is a massive victory for LoomRush! \par \vspace{1em}

32:00 - LoomRush's lone Knight attempted a Town Center raid on FeudalLord, but it had no impact. FeudalLord is booming with a fourth TC and more villagers. LoomRush continues to mass powerful Knights and Cavalry Archers, pushing forward! \par \vspace{1em}

32:27 - What a brutal engagement! LoomRush's Knights and Cavalry Archers just crushed FeudalLord's army at his forward stone mine. FeudalLord lost 27 units, including Eagles, Pikemen, and Monks, to LoomRush's mere 9 casualties. What a devastating loss for FeudalLord! \par \vspace{1em}

32:31 - A brief skirmish near LoomRush's base! FeudalLord loses a builder to a Knight, while LoomRush also loses a Knight. LoomRush's powerful Knights and Cavalry Archers continue to threaten FeudalLord's booming economy! \par \vspace{1em}

35:16 - LoomRush’s Cavalry Archers just stormed FeudalLord’s base, taking out an Eagle Warrior! But FeudalLord's defense was sharp, eliminating three of LoomRush's Cavalry Archers in a fierce exchange. What a hold! \par \vspace{1em}

37:59 - FeudalLord has completed a Castle near his forward secondary Stone Mine! This provides crucial defense and map control, protecting his vital resources. A big moment! \par \vspace{1em}

36:48 - A major battle erupted near LoomRush's base! FeudalLord's Eagle Warriors and Pikemen were completely wiped out, losing 30 units. LoomRush's Knights and Man-at-Arms lost only 10 units! A crushing defeat for FeudalLord! \par \vspace{1em}

38:28 - LoomRush, the Frankish player, just completed a new Town Center near his tertiary gold mine! This major economic expansion will fuel his powerful military and increase his villager lead! \par \vspace{1em}

37:24 - WHAT A CLASH! FeudalLord's Eagle Warriors and Pikemen were CRUSHED near LoomRush's Town Center, losing 27 units! LoomRush's military held strong, but he lost 14 villagers in the fight. A huge blow for FeudalLord! \par \vspace{1em}

39:26 - FeudalLord completes a Castle right by his starting Town Center! This powerful defensive structure will protect his economy and provide a safe retreat for his army. A solid move from FeudalLord! \par \vspace{1em}

38:45 - LoomRush's Cavalry Archers and Knights attacked FeudalLord's forward Stone Mine! Despite FeudalLord's Castle defense, LoomRush managed to pick off a crucial builder with no losses. A tough blow for FeudalLord's economy! \par \vspace{1em}

40:13 - FeudalLord's Eagle Warriors were completely crushed near LoomRush's Town Center! All 13 Eagles were lost. LoomRush, despite losing multiple villagers, held his ground. A decisive victory for LoomRush! \par \vspace{1em}

41:26 - A huge battle erupted near FeudalLord's base! FeudalLord lost 30 units, including his powerful Eagle and Jaguar Warriors. LoomRush lost even more, 39 units, including Knights and Long Swordsmen. FeudalLord is now advancing to Imperial Age! \par \vspace{1em}

41:55 - A brutal battle at LoomRush's forward gold mine! FeudalLord's Eagles were wiped out, but LoomRush lost 3 villagers and 4 Long Swordsmen! What a trade, as FeudalLord continues his Imperial Age research! \par \vspace{1em}

43:12 - A skirmish just concluded! FeudalLord's Pikeman was lost, but LoomRush's Long Swordsmen took no casualties. FeudalLord continues to push for Imperial Age, while LoomRush holds strong in Castle Age! \par \vspace{1em}

44:16 - What a massive battle in the west! FeudalLord's Jaguar and Eagle Warriors just crushed LoomRush's Cavalry Archers and Long Swordsmen. FeudalLord lost only 5 units, while LoomRush suffered a devastating 16 casualties! FeudalLord is nearing Imperial Age! \par \vspace{1em}

44:16 - And that's it! FeudalLord takes the win! What a comeback from FeudalLord, securing victory in a hard-fought battle! Congratulations to FeudalLord! \par \vspace{1em}

\subsection{Event Structuring}
\textbf{FeudalLord (Aztecs, blue) vs LoomRush (Franks, red)}

0:05 - Welcome to Arabia! On one side, we have FeudalLord, playing as the blue Aztecs. Their opponent, the red Franks, is LoomRush. Both players are starting strong with 5 villagers and a scout. \par \vspace{1em}

0:05 - FeudalLord, the blue Aztecs, has their main gold close and forward, but their main stone is also forward. LoomRush, the red Franks, mirrors this with a forward main gold and main stone. This could lead to intense early aggression! \par \vspace{1em}

10:29 - A fierce skirmish erupts near LoomRush's base! FeudalLord's aggressive Man-at-Arms push has paid off, taking down a red Farmer. FeudalLord maintains their military pressure with three Man-at-Arms deep, while LoomRush is now down a villager and relying on Scouts to defend! \par \vspace{1em}

11:59 - The battle on LoomRush's hill is over! LoomRush's Scouts have absolutely cleaned up FeudalLord's Man-at-Arms. All three of FeudalLord's aggressive Man-at-Arms are down, with LoomRush losing absolutely nothing! That's a huge defensive victory for the red Franks, completely repelling the blue Aztec's early aggression. \par \vspace{1em}

15:06 - Another major engagement has concluded near LoomRush's base! FeudalLord's attack, consisting of two Spearmen and three Archers, has been completely wiped out. LoomRush's defensive force, combining Skirmishers and Scout Cavalry, only lost one Scout Cavalry in return. LoomRush decisively defends their base again! \par \vspace{1em}

20:46 - A huge battle near FeudalLord's base has wrapped up! Both sides suffered massive losses, with FeudalLord losing 5 Archers and 2 Eagle Scouts. LoomRush also took heavy hits, losing 4 Skirmishers, 2 Scout Cavalry, and an Archer. It was an even trade of military units, but LoomRush is already in the Castle Age and still has Knights on the field! \par \vspace{1em}

25:59 - The brutal fight on FeudalLord's forward hill is over! LoomRush has taken a clear victory here, only losing two Skirmishers. FeudalLord suffered heavy losses, with two Eagle Scouts and two Builders going down. LoomRush's strong Knight and Cavalry Archer force proved superior in that engagement! \par \vspace{1em}

28:55 - A major battle near FeudalLord's base has concluded! FeudalLord's Monks, Pikemen, and Cavalry Archers went down. LoomRush only lost a Scout Cavalry and a Knight, securing a very favorable trade! Meanwhile, a massive battle is ongoing at FeudalLord's forward Stone Mine, with blue Eagle Warriors and Pikemen clashing against red Knights and Cavalry Archers! \par \vspace{1em}

32:31 - The brutal fight at FeudalLord's forward Stone Mine has wrapped up! LoomRush emerges victorious, losing only 9 units to FeudalLord's staggering 27 casualties, including many Eagle Warriors and Pikemen. Additionally, a skirmish concluded near LoomRush's base, where FeudalLord lost a Knight but took down a red Builder. The military counts are now even at 42 units each! \par \vspace{1em}

39:26 - What a bloodbath! LoomRush has absolutely decimated FeudalLord's army in back-to-back major engagements near their base, taking down nearly 60 military units while losing far fewer. LoomRush also lost some villagers, but it was a massive military win! FeudalLord is quickly reacting, completing two Castles, one forward and one at home. LoomRush, meanwhile, has completed a new Town Center. And now, a fierce battle rages near LoomRush's main Town Center! \par \vspace{1em}

44:16 - The fierce battle near LoomRush's main Town Center has ended! FeudalLord lost all 13 Eagle Warriors, but LoomRush paid a heavy price, losing 8 Farmers and 5 Long Swordsmen. A subsequent major clash near FeudalLord's base also saw massive casualties for both. But now it's over! FeudalLord has secured a huge victory in the west, trading just 5 units for 16 of LoomRush's! With that, FeudalLord wins the game! \par \vspace{1em}

\subsection{Expert Insights \& Event Structuring}
\textbf{FeudalLord (Aztecs, blue) vs LoomRush (Franks, red)}

0:05 - Welcome to Arabia! On the blue side, it's FeudalLord, the clever strategist playing Aztecs, starting with an Eagle Scout and extra gold for their fast-producing military. Opposing him is LoomRush, the red Franks, known for strategic depth, wielding powerful cavalry and a robust food economy with free farm upgrades. Let the game begin! \par \vspace{1em}

0:05 - Alright, taking a look at the map for both players! FeudalLord, as the Aztecs, has a very forward main gold and stone, forcing an aggressive posture early on. LoomRush's Franks also find their main resources out front. This is setting up for an intense, early clash for map control! \par \vspace{1em}

10:29 - FeudalLord's Man-at-Arms rush is in full swing, with three of them making their way towards LoomRush's base! And they've found their mark! A skirmish erupts near LoomRush's woodline, and FeudalLord's Man-at-Arms take down a Farmer! First villager down for LoomRush, a crucial early economic blow. Meanwhile, LoomRush is going for Scouts, with three out on the map. \par \vspace{1em}

13:05 - The Man-at-Arms push from FeudalLord has been completely shut down! LoomRush's five Scout Cavalry made quick work of all three Man-at-Arms on a forward hill, taking no losses themselves. This is a huge setback for FeudalLord, who now seems to be transitioning into Archer production, perhaps looking for a new angle of attack. \par \vspace{1em}

15:06 - An absolute bloodbath near LoomRush's base! FeudalLord's Feudal Age aggression is completely crushed. His entire force of 3 Archers and 2 Spearmen has been wiped out! LoomRush's Scout Cavalry and Skirmishers made short work of them, only losing a single Scout in return. FeudalLord is now left with almost no military and a significant setback! \par \vspace{1em}

20:46 - A huge engagement just wrapped up near FeudalLord's base! FeudalLord's Archer and Eagle Scout force was wiped out, losing 7 units. LoomRush's well-managed cavalry and skirmishers, despite taking losses, completely cleared the field. Crucially, LoomRush is now in Castle Age with Knights, while FeudalLord is still researching the age-up. A decisive victory for LoomRush! \par \vspace{1em}

25:59 - Another major battle just concluded on FeudalLord's forward hill! FeudalLord lost two Eagle Scouts and two Builders, while LoomRush only sacrificed two Skirmishers. A very cost-efficient exchange for LoomRush! FeudalLord has now hit Castle Age and even dropped two more Town Centers, trying to boost his economy, but LoomRush is pressing hard with a strong Knight and Cavalry Archer force! \par \vspace{1em}

28:55 - Another major engagement near FeudalLord's base just ended, and it was a disaster for the Aztecs! FeudalLord lost 7 units, including 4 Monks, while LoomRush only lost two! That's a huge economic blow. Now, an intense battle is still raging at FeudalLord's forward stone mine, with FeudalLord's Eagles and Pikemen clashing against LoomRush's Knights and Cavalry Archers! \par \vspace{1em}

32:31 - That massive engagement at FeudalLord's forward stone mine has concluded! It was devastating for FeudalLord, who lost 27 units, primarily Eagles and Pikemen. LoomRush, in contrast, only sacrificed 9 units, mostly Cavalry Archers and Knights, at that location. This has significantly weakened FeudalLord's military and bolstered LoomRush's already strong position. A quick trade saw FeudalLord lose a Knight but get a LoomRush Builder. \par \vspace{1em}

39:26 - FeudalLord's aggressive pushes have been brutally repelled! Two massive battles near LoomRush's base saw FeudalLord lose a combined 57 military units to LoomRush's 24, which included many villagers. In response, FeudalLord drops two Castles for defense. But LoomRush secures a new Town Center for economic growth, while FeudalLord's Eagles still pressure LoomRush's main Town Center! \par \vspace{1em}

44:16 - The pressure on LoomRush continues! That massive attack on LoomRush's main Town Center just concluded, FeudalLord's Eagles were wiped out but they took down 8 Farmers and 5 Long Swordsmen in the process. FeudalLord then decisively wins a huge engagement in the west, trading 5 of his units for 16 of LoomRush's, including many Cavalry Archers! With FeudalLord now researching Imperial Age, and LoomRush's army shattered, FeudalLord takes the win! FeudalLord is victorious! \par \vspace{1em}

\section{TaToH (IcTiTaN) vs Nicov (WoolyScout)}
\label{appendix:tatoh-nicov}
\subsection{Base}
\textbf{IcTiTaN (Vietnamese, blue) vs WoolyScout (Britons, red)}

0:05 - Welcome to Arabia! In this exciting match, we have IcTiTaN, playing as the blue Vietnamese, going up against WoolyScout, who is the red Britons. Let's get right into the action! \par \vspace{1em}

0:05 - Both IcTiTaN and WoolyScout have their main gold and stone near their Town Centers. Woodlines are also plentiful and well-placed, setting up solid early economies for both. \par \vspace{1em}

11:03 - IcTiTaN's Scout Cavalry probed WoolyScout's base, skirmishing with villagers! No losses occurred. Both players are building up their economies, now with 25 villagers each, and adding early military units. \par \vspace{1em}

14:52 - What a clash! An Archer from IcTiTaN has just defended his base, wiping out two of WoolyScout's Militia near IcTiTaN's territory! No losses for IcTiTaN in that skirmish! \par \vspace{1em}

15:09 - A major battle just erupted near WoolyScout's base! IcTiTaN's forces pushed in, losing a Spearman and a Scout Cavalry. But they took down one of WoolyScout's Farmers in a significant economic hit! \par \vspace{1em}

16:50 - A fierce skirmish at WoolyScout's base! IcTiTaN's attack was repelled, losing 2 Archers and a Skirmisher. WoolyScout's defenses held strong with no losses! Both players are now clicking up to Castle Age. \par \vspace{1em}

18:41 - IcTiTaN just launched a devastating raid on WoolyScout's economy! Two villagers, including a Lumberjack, were lost for WoolyScout near his base, with no losses for IcTiTaN's attacking force! \par \vspace{1em}

20:54 - A massive battle raged near WoolyScout's base! IcTiTaN's forces lost two units but completely wiped out WoolyScout's five Elite Skirmishers! A decisive victory for IcTiTaN in that engagement! \par \vspace{1em}

21:19 - A massive battle just concluded in the south-west! IcTiTaN lost 4 units, including Elite Skirmishers and a Crossbowman. WoolyScout took heavier military losses, losing 5 Crossbowmen in the fight! \par \vspace{1em}

22:11 - A skirmish at WoolyScout's Town Center just ended! IcTiTaN's attack was repelled, losing three units, including Elite Skirmishers and a Crossbowman. WoolyScout defended well, taking no losses! \par \vspace{1em}

25:08 - A major battle concluded near IcTiTaN's base! IcTiTaN lost one Elite Skirmisher. WoolyScout suffered heavy losses, with four Crossbowmen eliminated in the fight! A clear win for IcTiTaN! \par \vspace{1em}

29:07 - A massive battle just concluded near IcTiTaN's base! IcTiTaN lost 10 military units and 7 villagers. But WoolyScout's entire attacking force was wiped out, losing 18 military units in a devastating defeat! \par \vspace{1em}

31:25 - A skirmish just erupted at IcTiTaN's gold mine! IcTiTaN's forces, with Archers and a Mangonel, wiped out WoolyScout's Knight. No losses for IcTiTaN in that defensive victory! \par \vspace{1em}

32:12 - Another fierce battle near WoolyScout's forward position! IcTiTaN's forces wiped out two of WoolyScout's Lumberjacks with no losses! A devastating economic blow for WoolyScout! \par \vspace{1em}

33:21 - A massive battle just erupted in the north! IcTiTaN's entire attacking force, including 12 Crossbowmen and 5 Elite Skirmishers, was wiped out! WoolyScout suffered no losses in a dominant victory! \par \vspace{1em}

34:37 - A fierce battle concluded near IcTiTaN's base! IcTiTaN lost 5 units, including 4 villagers. But WoolyScout's entire attacking force of 3 Knights was wiped out! \par \vspace{1em}

36:51 - WoolyScout just completed a new Town Center in the north! That's a huge economic boost for the red Britons, expanding their reach and villager production! \par \vspace{1em}

38:01 - IcTiTaN just completed a new Town Center in the south-west! That's a big economic boost for the blue Vietnamese, expanding their villager production and map control. \par \vspace{1em}

39:33 - A brutal battle near IcTiTaN's base! IcTiTaN lost 49 units, including villagers and military. WoolyScout lost 28 military units, including Knights and Skirmishers. Heavy losses for both! \par \vspace{1em}

39:33 - And that's it! WoolyScout has emerged victorious! The red Britons take the win against IcTiTaN in a hard-fought match. Congratulations to WoolyScout! What an intense game! \par \vspace{1em}

\subsection{Expert Insights}
\textbf{IcTiTaN (Vietnamese, blue) vs WoolyScout (Britons, red)}

0:05 - Welcome to Arabia! We have IcTiTaN, the creative strategist from Iron Phalanx, playing as the blue Vietnamese, known for their strong Archery Range units and economic bonuses. Up against him is WoolyScout, the macro master of Wololo Warriors, commanding the red Britons, with their iconic Longbowmen and faster shepherds. Let's see this unfold! \par \vspace{1em}

0:05 - Both players face forward main gold veins, meaning early pressure or strong defenses will be key! IcTiTaN's Vietnamese and WoolyScout's Britons will need to secure those vital resources quickly to power their economies. \par \vspace{1em}

7:05 - WoolyScout is going for an early Drush! At 7 minutes, we're seeing some aggressive intent from the Briton player. They'll be looking to disrupt IcTiTaN's economy and delay his build with those initial Militia. This could be a rocky start for the Vietnamese! \par \vspace{1em}

13:05 - IcTiTaN is confirming an Archery Range strategy at 13 minutes! With the Vietnamese bonuses, those Archers will be very strong. Expect to see some heavy ranged pressure soon from the blue player! \par \vspace{1em}

11:03 - IcTiTaN's scout found villagers at WoolyScout's base, but no kills. Crucially, IcTiTaN now has a Skirmisher out, ready to defend against that ongoing Drush! Both players are neck and neck in villager count. \par \vspace{1em}

14:05 - Now WoolyScout is also committing to Archers at 14 minutes! This sets up a crucial ranged battle. With their Briton bonuses, WoolyScout's Archers will have a strong edge in this developing mirror match-up! \par \vspace{1em}

14:52 - IcTiTaN's defense pays off! An Archer near his base just wiped out WoolyScout's two Militia, ending that early Drush! IcTiTaN has defended perfectly and taken the first kills of the game! \par \vspace{1em}

15:09 - IcTiTaN is raiding WoolyScout's base! A fierce skirmish erupts, and IcTiTaN picks off a Briton villager! But the Vietnamese raiders suffer casualties, losing a Spearman and Scout. WoolyScout has held on! \par \vspace{1em}

16:50 - A terrible raid for IcTiTaN! Near WoolyScout's base, his Archers and Skirmisher were completely wiped out. Three units lost for IcTiTaN, zero for WoolyScout! IcTiTaN's aggressive play backfired, a major setback in this archery war. \par \vspace{1em}

18:41 - Correction: IcTiTaN's raid on WoolyScout was a success! His Archers and Skirmishers wiped out two villagers at WoolyScout's base, taking zero losses! That's a huge economic blow for the Britons! \par \vspace{1em}

20:54 - What a clash! Near WoolyScout's base, IcTiTaN's Crossbowmen and Elite Skirmishers just annihilated WoolyScout's entire Skirmisher force! IcTiTaN lost two units but completely wiped out five of WoolyScout's! A massive military win for IcTiTaN! \par \vspace{1em}

21:19 - Another intense engagement in the south-west! IcTiTaN's Elite Skirmishers and Crossbowmen clashed with WoolyScout's Crossbowmen, Scout, and Knight. WoolyScout lost five Crossbowmen, while IcTiTaN lost four units. A very close fight! \par \vspace{1em}

22:11 - A disaster for IcTiTaN! His Elite Skirmishers and Crossbowman were wiped out attacking WoolyScout's base. WoolyScout's Knight and Builder survived completely unscathed! This is a huge defensive victory for WoolyScout! \par \vspace{1em}

25:08 - IcTiTaN defends brilliantly again! Near his base, his Elite Skirmishers and Crossbowmen wiped out four of WoolyScout's Crossbowmen, losing only one unit! A decisive win for IcTiTaN! \par \vspace{1em}

29:07 - A colossal battle at IcTiTaN's base! WoolyScout's powerful force, including 12 Knights, was obliterated, losing 20 units! IcTiTaN took 18 casualties, but his defense was decisive, crippling WoolyScout's military. \par \vspace{1em}

31:25 - IcTiTaN's defense is impenetrable! At his forward gold, his Crossbowmen and Mangonel swiftly eliminated a lone Knight from WoolyScout without taking any losses! A perfect hold for IcTiTaN! \par \vspace{1em}

32:12 - IcTiTaN launches a devastating raid at WoolyScout's forward position! His Crossbowmen and Elite Skirmishers eliminated two Lumberjacks, while suffering zero losses! That's a huge economic blow for WoolyScout, and a perfectly executed attack by IcTiTaN! \par \vspace{1em}

33:21 - ABSOLUTE DESTRUCTION! In the north, WoolyScout's Knights and Mangonels just wiped out IcTiTaN's entire force of 17 Crossbowmen and Skirmishers without taking a single loss! A crushing victory for WoolyScout! \par \vspace{1em}

34:37 - A tense battle erupts near IcTiTaN's base! WoolyScout's three Knights were all eliminated, but IcTiTaN paid a heavy price, losing four villagers and an Elite Skirmisher! A military win for IcTiTaN, but costly. \par \vspace{1em}

36:51 - WoolyScout has just completed a new Town Center in the north! This is a massive economic play, looking to skyrocket his villager production and gain a vital lead in the game! \par \vspace{1em}

38:01 - IcTiTaN has just completed a new Town Center in the south-west! This is a huge economic play for the Vietnamese player, looking to boost his villager production and secure a strong economic lead in the late game! \par \vspace{1em}

39:33 - A massive clash at IcTiTaN's base! WoolyScout's attacking force, including many Knights, was largely wiped out. IcTiTaN suffered heavy losses of both military and economic units, but successfully defended against the large raid! \par \vspace{1em}

39:33 - And that's it! WoolyScout has secured the victory! What a performance by the Briton player, ultimately overcoming IcTiTaN in this intense matchup! Congratulations to WoolyScout! \par \vspace{1em}

\subsection{Event Structuring}
\textbf{IcTiTaN (Vietnamese, blue) vs WoolyScout (Britons, red)}

0:05 - Welcome to Arabia! On one side, we have IcTiTaN, playing as the Blue Vietnamese. Facing him is WoolyScout, commanding the Red Britons. Both players begin with five villagers, focusing on food, and one scout cavalry to explore their surroundings. \par \vspace{1em}

0:05 - IcTiTaN's map reveals a forward main gold pile of seven tiles and two forward stone piles. His woodlines are also mostly forward. For WoolyScout, the main gold is also a forward seven-tile pile, with two stone piles situated forward. Woodlines appear plentiful and forward for both players. \par \vspace{1em}

11:03 - Both players are now firmly in the Feudal Age, with both IcTiTaN and WoolyScout having 25 villagers. IcTiTaN's Scout Cavalry made an aggressive move into WoolyScout's base, attempting to raid villagers, but WoolyScout's economy remains unharmed. Both players are now building Archery Ranges, setting the stage for more ranged unit engagements! \par \vspace{1em}

16:50 - Action erupts! IcTiTaN's Archer took down two Militia near his base. Then, IcTiTaN pushed aggressively into WoolyScout's lumberline, losing a Spearman and Scout but picking off a Farmer. However, WoolyScout responded effectively, wiping out three of IcTiTaN's Archers and Skirmisher in another skirmish! Both players are now pushing for Castle Age, with WoolyScout holding a military advantage. \par \vspace{1em}

21:19 - Both players are now firmly in Castle Age, and the battles are heating up! IcTiTaN's archer units raided WoolyScout's base, eliminating two of his villagers. In the south-west, a major clash erupted as WoolyScout's Crossbowmen and a Knight faced off against IcTiTaN's Elite Skirmishers. WoolyScout lost five Crossbowmen, while IcTiTaN lost three Elite Skirmishers and a Crossbowman. Elsewhere, IcTiTaN's forces eliminated five more of WoolyScout's Elite Skirmishers! \par \vspace{1em}

22:11 - Big action around the map! Near WoolyScout's Town Center, IcTiTaN's Elite Skirmishers and a Crossbowman were completely wiped out by a lone Knight and Builder defending, a massive loss for IcTiTaN! Simultaneously, WoolyScout's Crossbowmen are now aggressively raiding IcTiTaN's base, targeting his villagers! WoolyScout has also pushed out a Mangonel, while IcTiTaN is researching Leather Archer Armor. \par \vspace{1em}

25:08 - Massive fights continue! Near IcTiTaN's base, a fierce clash just ended where IcTiTaN's Elite Skirmishers and Crossbowmen successfully defended, taking down four of WoolyScout's Crossbowmen while losing only one Elite Skirmisher! But the pressure isn't over! WoolyScout is now pushing hard into IcTiTaN's base with a powerful force of Knights, Crossbowmen, and a Mangonel, threatening the Town Center and villagers! IcTiTaN is fighting back with Elite Skirmishers and Crossbowmen of his own. \par \vspace{1em}

32:12 - That massive raid on IcTiTaN's base just ended! Both sides took heavy losses, but WoolyScout's Knights and Crossbowmen were completely wiped out, losing 20 units! IcTiTaN lost 18 units, including many villagers. Now, IcTiTaN is on the offensive! His archers just eliminated a Knight near his gold, and then pushed into WoolyScout's lumberline, taking down two Lumberjacks. IcTiTaN is turning the tables! \par \vspace{1em}

38:01 - WoolyScout landed a decisive blow, wiping out 17 of IcTiTaN's units in the north! But IcTiTaN retaliated, destroying three Knights near his own base, defending against the raid! Now, Mangonels are clashing near IcTiTaN's Town Center, and a big fight is underway at WoolyScout's gold mine, as IcTiTaN's Crossbowmen face Knights and Skirmishers! Both players are also booming with new Town Centers! \par \vspace{1em}

39:33 - The massive raid on IcTiTaN's base just concluded! IcTiTaN suffered devastating losses, 49 units mostly villagers and military. WoolyScout lost 28 military units, but his persistent pressure secured a decisive advantage. Both expanded with new Town Centers, but WoolyScout's sustained economic and military lead sealed the game. WoolyScout takes the win! \par \vspace{1em}

\subsection{Expert Insights \& Event Structuring}
\textbf{IcTiTaN (Vietnamese, blue) vs WoolyScout (Britons, red)}

0:05 - Welcome to Arabia! In the blue corner, we have IcTiTaN, known for his creative strategies, commanding the Vietnamese with their unique map vision. Opposite him in red, it's WoolyScout, renowned for strong macro play, leading the Britons and their powerful, long-range archers. An archer showdown is brewing! \par \vspace{1em}

0:05 - Looking at the map, both players have very forward main gold and stone! IcTiTaN’s key resources are exposed, but WoolyScout’s seem even more so. This forward placement for the Briton player could force early aggression to secure them, or demand significant early walling and defensive towers to protect those vital resources. A tough map! \par \vspace{1em}

7:05 - At just over 7 minutes, we're seeing an early aggressive move! WoolyScout, the Briton player, is going for a Dark Age rush! This Drush means we'll likely see early militia or scouts harassing IcTiTaN. Given the forward resources, WoolyScout could be looking to apply immediate pressure and deny key gold or stone, forcing IcTiTaN onto the defensive right from the start! What a bold play! \par \vspace{1em}

13:05 - The Drush is on! WoolyScout has those two Militia pushing forward, looking to apply early pressure on IcTiTaN. But IcTiTaN isn't sitting back! He's now in Feudal Age, built an Archery Range, and has a Skirmisher out, ready to defend against that Militia aggression. Both players are neck and neck with 25 villagers, but it's a military showdown now in the Feudal Age! \par \vspace{1em}

16:50 - The early Drush from WoolyScout fizzles out, with IcTiTaN's Archer taking down both Militia! But the pressure is still on, as IcTiTaN pushes back, raiding WoolyScout's base! IcTiTaN loses a Spearman and Scout, but WoolyScout loses a farmer in a costly defense! Then, WoolyScout decisively defends another push, wiping out IcTiTaN's small army of Archers and Skirmishers! Both players are now clicking up to Castle Age, with WoolyScout gaining a military lead! \par \vspace{1em}

21:19 - Both players are now in the Castle Age! IcTiTaN's aggression continues, first winning a decisive skirmish near WoolyScout's base, eliminating several Elite Skirmishers. Then, a major battle erupts in the south-west! IcTiTaN's combined Elite Skirmishers and Crossbowmen bravely face WoolyScout's Crossbowmen and Knights. IcTiTaN takes the military advantage, inflicting more casualties and maintaining a villager lead, pressuring WoolyScout's economy! \par \vspace{1em}

22:11 - IcTiTaN's aggressive push at WoolyScout's Town Center was crushed! IcTiTaN lost two Elite Skirmishers and a Crossbowman for no losses to WoolyScout. Now, WoolyScout is hitting back hard, raiding IcTiTaN's base with Crossbowmen, forcing villagers off work! IcTiTaN has a slight villager advantage, but WoolyScout just produced a Mangonel to go with his Crossbowmen and Knights. Intense action! \par \vspace{1em}

25:08 - IcTiTaN secures a key defensive win near his base, taking down four Crossbowmen from WoolyScout for just one Elite Skirmisher! But WoolyScout's raid continues unabated! He's pressing hard with Crossbowmen, Knights, and that threatening Mangonel right on IcTiTaN's Town Center. IcTiTaN holds a villager lead of 47 to 44 and has brought out his own Mangonel. The pressure is immense! \par \vspace{1em}

32:12 - Huge, sustained battles near IcTiTaN's Town Center! WoolyScout's large force of Knights, Crossbowmen, and Mangonels were decimated! IcTiTaN defended valiantly, losing some villagers, but now he's pushing back! He wiped out a lone Knight at his gold and continues to raid WoolyScout's forward lumber camp, taking down two more villagers! IcTiTaN now leads in military units and applies immense pressure! \par \vspace{1em}

38:01 - A devastating battle in the north! WoolyScout completely wiped out IcTiTaN's entire Crossbowman and Skirmisher army, suffering no losses himself – a massive military swing! IcTiTaN successfully defended a Knight raid at his Town Center. Both players are expanding their economies with new Town Centers. Now, WoolyScout pushes hard, initiating a Mangonel vs Mangonel duel at IcTiTaN's base, while IcTiTaN's Crossbowmen raid WoolyScout's gold! \par \vspace{1em}

39:33 - What an incredible end to a back-and-forth game! The massive raid near IcTiTaN's base finally concludes! IcTiTaN suffers devastating losses, with 49 units, including many villagers, falling to WoolyScout's assault. WoolyScout took significant damage too, losing 28 units, but the sheer economic damage inflicted on IcTiTaN was too much to overcome! WoolyScout has secured the victory! GG! \par \vspace{1em}

\section{TheViper (CastleDropper) vs Jordan (WallMaster77)}
\label{appendix:viper-jordan}
\subsection{Base}
\textbf{CastleDropper (Bulgarians, yellow) vs WallMaster77 (Indians, blue)}

0:04 - Welcome to Arabia! We have CastleDropper, playing as the yellow Bulgarians, facing off against WallMaster77, the blue Indians. Both players start their game with five villagers and one scout cavalry. It's going to be an exciting match here today! \par \vspace{1em}

0:04 - Let's look at the map! CastleDropper has a good main gold and stone close to their Town Center, with multiple woodlines. WallMaster77 also has close main gold and stone, and solid woodlines. Both players have a decent resource spread. \par \vspace{1em}

10:33 - CastleDropper is making an early aggressive move! They are placing a forward Watch Tower right near WallMaster77's base. This could put immense pressure on WallMaster77 very early on! What a start! \par \vspace{1em}

11:18 - That Watch Tower is now complete! CastleDropper's forward Watch Tower is up and putting immediate pressure on WallMaster77's base. This early aggression from the Bulgarians could dictate the pace of this game! \par \vspace{1em}

11:20 - WallMaster77 responds to the pressure! The blue Indians have completed their own Watch Tower, placed defensively near their starting Town Center. They're looking to fortify their position against CastleDropper's early aggression! \par \vspace{1em}

11:40 - CastleDropper is relentless! Another forward Watch Tower foundation is down near WallMaster77's base. The yellow Bulgarians are doubling down on this early pressure! \par \vspace{1em}

12:24 - And just like that, CastleDropper's second forward Watch Tower is complete! The yellow Bulgarians are relentless with this double tower aggression, locking down WallMaster77's base! \par \vspace{1em}

12:10 - A raid just concluded near WallMaster77's base, but both sides managed to avoid any casualties! CastleDropper now has two forward Watch Towers. WallMaster77 is responding by training Archers and a Spearman to defend against this tower rush! \par \vspace{1em}

15:05 - CastleDropper is going for a THIRD forward Watch Tower! This one is being placed right next to WallMaster77's main gold mine, denying crucial resources! \par \vspace{1em}

15:06 - CastleDropper completes yet another Watch Tower! This one is purely defensive, placed near their own main gold mine, not a forward push! \par \vspace{1em}

16:27 - A huge clash! CastleDropper's Man-at-Arms and Skirmishers took heavy losses, with 9 military units and a Villager falling to WallMaster77's Archers and Scout Cavalry near the gold! WallMaster77 only lost two Archers! \par \vspace{1em}

16:27 - And that's the game! WallMaster77, the blue Indians, emerge victorious! Despite the intense early pressure from CastleDropper's tower rush, WallMaster77 held strong and secures the win! What a comeback! \par \vspace{1em}

\subsection{Expert Insights}
\textbf{CastleDropper (Bulgarians, yellow) vs WallMaster77 (Indians, blue)}

0:04 - Welcome to Arabia! In the yellow corner, CastleDropper, renowned for his versatility, commands the Bulgarians. Facing him in blue is WallMaster77, known for his strong macro and strategic depth, playing the Indians. It's an exciting matchup between two skilled players! \par \vspace{1em}

0:04 - Both players face forward main gold and stone. This map setup means early aggression is likely, as these vital resources are vulnerable! CastleDropper's Bulgarians could use their forward stone for an early tower push. \par \vspace{1em}

7:04 - CastleDropper is indeed going for a Drush! He's committing to early aggression with militia, hoping to disrupt WallMaster77's economy and slow down his villager production. This is an aggressive start! \par \vspace{1em}

10:04 - CastleDropper is indeed transitioning from his Drush into a full-blown Tower Rush! He's got a forward tower going up, putting immense pressure on WallMaster77's economic operations. This is a very aggressive opening! \par \vspace{1em}

10:33 - CastleDropper is relentless! Another forward Watch Tower is going up near WallMaster77's base. This is immense pressure, deeply threatening WallMaster77's resource collection and forcing a quick defensive reaction! \par \vspace{1em}

11:04 - CastleDropper is indeed shifting strategies, now identified with a Man-at-Arms rush! This continues his aggressive feudal age, as he upgrades his militia or produces new Man-at-Arms. The pressure on WallMaster77 intensifies! \par \vspace{1em}

11:18 - Another Watch Tower for CastleDropper is now complete right near WallMaster77's base! The Bulgarian aggression is relentless, putting immense pressure on WallMaster77's economic operations. This is a very strong power play in the Feudal Age! \par \vspace{1em}

11:20 - WallMaster77 completes a Watch Tower! This is a much-needed defensive response, as he fortifies his base against CastleDropper's relentless aggression. He's trying to weather the storm! \par \vspace{1em}

11:40 - CastleDropper is laying the foundation for yet another Watch Tower right near WallMaster77's Town Center! The pressure from the Bulgarians is absolutely relentless, aiming to completely shut down WallMaster77's economy. \par \vspace{1em}

12:04 - Huge strategic shift! WallMaster77 is now going for Archers. This is a smart move against CastleDropper's infantry and much-needed ranged defense for those towers. He's adapting to the relentless pressure! \par \vspace{1em}

12:24 - Another Watch Tower for CastleDropper is now complete right by WallMaster77's Town Center! The Bulgarian aggression is absolutely relentless, putting immense pressure on WallMaster77's economy and villager production! This is intense! \par \vspace{1em}

12:10 - A protracted fight near WallMaster77's base concludes! CastleDropper's Man-at-Arms and towers faced WallMaster77's growing defense of Archers and Spearmen. No units lost for either side, a fierce but indecisive skirmish! \par \vspace{1em}

15:05 - CastleDropper is laying the foundation for yet another Watch Tower right on WallMaster77's forward, main Gold Mine! This is absolute economic strangulation if it goes up. \par \vspace{1em}

15:06 - CastleDropper's Watch Tower is now complete! It's defensively positioned near his own forward, main gold mine. He's securing his vital resources, perhaps consolidating his economy after all that aggression. \par \vspace{1em}

16:27 - A massive battle concludes at WallMaster77's gold mine! CastleDropper loses 5 Man-at-Arms, 4 Skirmishers, and a Villager! WallMaster77 defends, only losing 2 Archers! A huge defensive victory for WallMaster77! \par \vspace{1em}

16:27 - And that's it! WallMaster77 has emerged victorious, taking the win in this intense game! After all that early pressure, WallMaster77's strong defense and economic play sealed the deal! What a comeback! GG! \par \vspace{1em}

\subsection{Event Structuring}
\textbf{CastleDropper (Bulgarians, yellow) vs WallMaster77 (Indians, blue)}

0:04 - Welcome to Arabia! In yellow, CastleDropper commands the Bulgarians. Their opponent, WallMaster77, is blue, leading the Indians. Both are set for a showdown on this open map! \par \vspace{1em}

0:04 - For CastleDropper, their main gold is forward, with two other gold piles also forward. Their main stone is forward, plus another stone pile. All their woodlines are also forward. WallMaster77 has a forward main gold, two other gold piles forward, and forward main and secondary stone. Their woodlines are all forward too! \par \vspace{1em}

12:24 - CastleDropper is pushing hard with forward Watch Towers! Their first tower is up, and they've been raiding WallMaster77's base with Man-at-Arms, though no units were lost in that skirmish. WallMaster77 has built a defensive tower at home. Now, CastleDropper has just completed a *second* forward Watch Tower, right on top of WallMaster77's Town Center! \par \vspace{1em}

16:27 - CastleDropper is pushing hard! They founded *another* forward Watch Tower at WallMaster77's main gold! But a huge raid just ended there! CastleDropper lost 9 military units and a Villager, while WallMaster77 only lost 2 Archers! That devastating engagement has sealed the deal: WallMaster77 takes the victory! \par \vspace{1em}

\subsection{Expert Insights \& Event Structuring}
\textbf{CastleDropper (Bulgarians, yellow) vs WallMaster77 (Indians, blue)}

0:04 - Welcome to Arabia! In the yellow corner, it's CastleDropper, known for his versatility, playing as the powerful Bulgarians with their unique Konniks and infantry. Facing him in blue is WallMaster77, celebrated for his strong macro and strategic depth, commanding the economic might of the Indians! \par \vspace{1em}

0:04 - Interesting map generation here! Both players have their main gold quite forward, meaning these crucial resources will be vulnerable early on. CastleDropper's Bulgarians, with their strong early game, might feel pressured to make an aggressive move to secure that gold. WallMaster77's Indians will need to wall up carefully or expect early pressure! \par \vspace{1em}

12:24 - And CastleDropper is truly living up to his name, dropping not one, but two forward Watch Towers deep into WallMaster77's territory! This aggressive Tower Rush, combined with his Man-at-Arms push, is putting immense pressure on WallMaster77's base. WallMaster77 is responding by building his own defensive Watch Tower and producing Archers, trying to hold off the Bulgarian onslaught! No casualties yet, but the fight is heating up! \par \vspace{1em}

16:27 - What a turn of events! That raid on WallMaster77's forward gold has concluded, and it's a disaster for CastleDropper! He lost 5 Man-at-Arms, 4 Skirmishers, and a Villager, while WallMaster77 only lost 2 Archers. That's a huge blow to the Bulgarian aggression. \par \vspace{1em}

Adding to the pressure, CastleDropper has finished a Watch Tower at his own forward gold, but it's too little, too late. WallMaster77's defense held strong, and it seems that crushing counter-attack was decisive... the game has ended! WallMaster77 takes the victory!
